### 2024-09-24 paper/paper.qmd
Completed draft 1 of paper, references to be added

---
title: "Analysis of Crime Victims by Type and Gender from 2014 to 2023"
author: 
  - Kevin Roe
thanks: "All data, R code, and other files are available in the following GitHub repository: <https://github.com/Kanghyunroe/victims_of_crimes/tree/master>"
date: today
date-format: long
abstract: "This paper analyzes data from the Police Annual Statistical Report – Victims of Crime for the years 2014 to 2023 in the City of Toronto. The paper reveals there are 2% more male victims overall, but women are disproportionately affected in certain crime categories: female victims outnumber males by 78 percentage points in sexual crimes and by 9 percentage points in other crimes. These differences show how vulnerability to specific crimes varies by sex. Further research should focus on victims of fatal crimes to explore these disparities."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| echo: false
#| warning: false
#### Preamble ####

# Purpose: Generates a paper titled "Analysis of Crime Victims by Type and gender from 2014 to 2023" in PDF format when rendered.
# Author: Kevin Roe
# Date: 24 September 2024
# Contact: kevin.roe@mail.utoronto.ca
# License: MIT
# Pre-requisites: Run the files "01-download_data.R" and "02-data_cleaning.R"
# Any other information needed? None.
```

```{r}
#| include: false
#| warning: false
#| message: false

#### Loading packages ####
#install.packages("scales")
#install.packages("tidyverse")
# You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.
# The remainder of this paper is structured as follows. @sec-mydatasection
library(tidyverse)
library(ggplot2)
library(scales)

cleaned_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction {#sec-introduction}

In every society, keeping crime at a minimum has been a critical priority. Notably, certain demographic groups have been more affected by crime than others. For example, in the United States in 2008, there were a greater number of female nonfatal violent domestic violence victims than male ones (Catalano et al.). Similarly, Toronto is no exception to this phenomenon. The City of Toronto notes that certain groups of people are more susceptible to Gender-based violence than others.

In this paper, I analyze the number of crime victims by sex in Toronto from 2014 to 2023. I categorize crime into four different areas: Assault, Sexual Violation, Robbery, and Other. Furthermore, assault is further broken down into different types of assault noted by the Toronto Police Department: assault on a peace officer, aggravated assault on a peace officer, resisting arrest, and assault on a peace officer with a weapon. We find that there are more female victims of sexual crimes and other broad crimes than men, there are more male victims for non-fatal crimes such as assault or robbery.

The remainder of the paper is structured as follows: Section 2 discusses the data; Section 3 presents the results; and Section 4 discusses our findings and highlights key weaknesses in the study.

# Data {#sec-data}

## Overview {#sec-data-overview}

This dataset, “Police Annual Statistical Report – Victims of Crime”, was published and refreshed on August 2nd, 2024, by the Toronto Police Services. The Toronto Police Service publishes various datasets on public safety to inform the public with information regarding public safety and self-awareness. Following the Municipal Freedom of Information and Protection of Privacy Act, the Toronto Police Service ensured to also protect the privacy of individuals involved in the reported crimes.  The dataset is updated annually, is open data, and can be used if an attribution statement @sec-appendix-attribution and is properly cited (Insert Citation).

The variables of interest in the paper are CrimeType (named “CATEGORY” in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robberty and Other; AssaultType (named “ASSAULT_SUBTYPE” in the original dataset), which specifies assault on peace officers into the subtypes noted in the Introduction (INSERT LINK); Sex (named “SEX” in the original dataset), which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count, which notes the number of identified victims who share the same demographic characteristics reported in the dataset. Other characteristics such as age group, reported year and age cohort were not included as the variables were beyond the scope of the paper.

The paper uses the R programming language (insert citation) to analyze the dataset. The 'tidyverse' package was used to simulate the dataset and generate tests. The 'tidyverse' (Citation) and 'opendatatoronto' (citation) packages were used to download the Victims of Crime dataset. Then, the 'tidyverse' package (citation) was to clean and test the raw dataset. Finally, 'ggplot2' (citation), 'tidyverse' (citation), and 'scales' (citation) packages were used to create data visualization figures to highlight the results.

## Results {#sec-data-results}

```{r}
#| include: false
#| warning: false
#| message: false

#### Load cleaned dataset ####

cleaned_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

After loading the dataset using the R programming language (citation), I used the 'tidyverse' (citation) package, and the 'ggplot2' (citation) package. The R code to do so was adapted from Alexander (insert citation).

```{r}
#| echo: false
#| warning: false

#### Organizing data for graph use ####
summary <- cleaned_data %>%
  group_by(Sex) %>%
  summarise(Victims_of_Crimes = sum(Count))

```

```{r}
#| label: fig-total
#| fig-cap: Victims of Crimes by Sex from 2014 to 2023
#| echo: false
#| warning: false

#### First Graph ####
ggplot(summary, aes(x = Sex, y = Victims_of_Crimes, fill = Sex)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Victims_of_Crimes),
            position = position_dodge(width = 0.9),
            vjust = -0.2,
            size = 3.5) +
  labs(x = "Sex",
       y = "Victims of Crimes") +
  scale_y_continuous(labels = comma) +
  theme_classic()

```

@fig-total shows that there are 137,527 male victims of crime from 2014 to 2023 and 134726 female ones, representing  a difference of 2,801 people or 2.1\%. @fig-total shows that there are more male crime victims than female ones, albeit the difference does not seem to be economically significant.

\newpage

```{r}
#| echo: false
#| warning: false

#### Organizing information for graphical use ####
summary_crime <- cleaned_data %>%
  group_by(CrimeType, Sex) %>%
  summarise(Victims_of_Crimes = sum(Count)) %>%
  ungroup() %>%
  mutate(CrimeType = fct_relevel(CrimeType, setdiff(unique(CrimeType), "Other"),
                                           "Other"))
```

```{r}
#| label: fig-crimetype
#| fig-cap: Breakdown of Different Categories of Crime by Gender from 2014 to 2023
#| echo: false
#| warning: false

#### Second Graph ####
ggplot(summary_crime, aes(x = CrimeType, y = Victims_of_Crimes, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = Victims_of_Crimes),
            position = position_dodge(width = 0.9),
            vjust = -0.2,
            size = 2.5) +
  labs(x = "Crime Type", 
       y = "Number of Victims") +
  scale_y_continuous(labels = comma) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust =1))

```

@fig-crimetype shows that while there are more assault and robbery male victims, the number of female victims for sexual crimes and other crimes remain significantly higher than male ones. Female victims outnumber males by 78 percentage points in sexual crimes and by 9 percentage points in other crimes.

\newpage

```{r}
#| echo: false
#| warning: false

#### Organizing information for graphical use ####
cleaned_assault <- cleaned_data %>%
  filter(AssaultType != "N/A")

summary_assault <- cleaned_assault %>%
  group_by(AssaultType, Sex) %>%
  summarise(Victims_of_Crimes = sum(Count)) %>%
  mutate(AssaultType = str_wrap(AssaultType, width = 15)) %>%
  ungroup() %>%
  mutate(AssaultType = fct_relevel(AssaultType, setdiff(unique(AssaultType), 
                                                        "Other"), "Other"))
```

```{r}
#| label: fig-assault
#| fig-cap: Victims of Crimes by Assault Type from 2014 to 2023
#| echo: false
#| warning: false

#### Third Graph ####
ggplot(summary_assault, aes(x = AssaultType, y = Victims_of_Crimes, 
                            fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = Victims_of_Crimes),
            position = position_dodge(width = 0.9),
            vjust = -0.2,
            size = 2.5) +
  labs(x = "Assault Type",
       y = "Number of Victims") +
  scale_y_continuous(labels = comma) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust =1))

```
@fig-assault shows that diving deeper into the different crimes under assault shows that there are more male assault victims than female ones  across all types of assault, including different types of assault to a peace officer.

\newpage

# Discussion

@fig-total showed that there were 2.1\% more reported male crime victims than female ones. However, when we break this down, the number of female victims outnumber males by 78 percentage points and 9 percentage points for sexual and other crimes, respectively. Further, when dissecting assault victims, we don’t see this same differential and there are more male victims than female ones.

Analyzing the data shows that vulnerability to specific crimes varies by sex. @fig-crimetype shows that women are more susceptible to gender-based violence such as sexual assault or harassment. According to Statistics Canada, approximately 4.7 million or 30\% of all women aged 15 or older have been sexually assaulted outside of an intimate relationship at least once. The data in figure 2 shows that from 2014 to 2023 there were 21,515 instances of sexual violence to women in the city of Toronto and only 3,181 for men, a disproportionate difference.

However, @fig-crimetype shows that men are also more susceptible to being victims of assault, and robbery  compared to women. @fig-crimetype shows that, across the past nine years, there were 11,169 more male assault victims than female. When broken into different categories, @fig-assault highlights that there are more male victims across all subsections of assault than female. Because assault is the most common category of crime in Toronto, the 2\% difference between male and female victims in @fig-total is fueled by the difference underlined in the assault category. Overall, the data analyzed today shows that both sexes are susceptible to different crimes. While males represent a greater proportion of total crime victims, women are disproportionately affected in sexual crimes.

When working with the dataset, there are a few limitations and next steps to keep in mind. First, the data may have some duplicates because the "Count"(INSERT CHANGE) variable could include the same person if they were victimized for the same offence during the same period. In addition, the dataset does not breakdown other into specific crimes under "CrimeType", making it difficult to analyze fatal victims of crime. Further research should look at if women or men are more susceptible to fatal crimes such as manslaughter or first-degree murder.

\newpage

\appendix

# Appendix {#sec-appendix}

## Dataset and Graph Sketches {#sec-appendix-sketches}

Sketches for the desired data set and the graphs are available in the GitHub Repository.

## Data Cleaning Methodology {#sec-appendix-cleaning}

Select columns from the raw data set were filtered out and data entries were renamed for simplicity.

## Data Attribution Statement {#sec-appendix-attributiojn}

"This data contains information licensed under the Open Government License - Toronto" (insert citation).

\newpage

# References (To be added) {#sec-references}

### 2024-09-24 paper/paper.qmd
added reference to R for initial submission

@@
 Notably, certain demographic groups have been more affected by crime than others.
-For example, in the United States in 2008, there were a greater number of female nonfatal violent domestic violence victims than male ones (Catalano et al.).
+For example, in the United States in 2008, there were a greater number of female nonfatal violent domestic violence victims than male ones (Catalano et al.
+INSERT STATS).
 Similarly, Toronto is no exception to this phenomenon.
@@
 
-The remainder of the paper is structured as follows: Section 2 discusses the data; Section 3 presents the results; and Section 4 discusses our findings and highlights key weaknesses in the study.
+The remainder of the paper is structured as follows: @sec-data-overview discusses the data; @sec-data-results presents the results; and @sec-discussion discusses our findings and highlights key weaknesses in the study.
 
@@
 
-The variables of interest in the paper are CrimeType (named “CATEGORY” in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robberty and Other; AssaultType (named “ASSAULT_SUBTYPE” in the original dataset), which specifies assault on peace officers into the subtypes noted in the Introduction (INSERT LINK); Sex (named “SEX” in the original dataset), which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count, which notes the number of identified victims who share the same demographic characteristics reported in the dataset.
+The variables of interest in the paper are CrimeType (named “CATEGORY” in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robberty and Other; AssaultType (named “ASSAULT_SUBTYPE” in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named “SEX” in the original dataset), which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count, which notes the number of identified victims who share the same demographic characteristics reported in the dataset.
 Other characteristics such as age group, reported year and age cohort were not included as the variables were beyond the scope of the paper.
 
-The paper uses the R programming language (insert citation) to analyze the dataset.
+The paper uses the R programming language [@citeR] to analyze the dataset.
 The 'tidyverse' package was used to simulate the dataset and generate tests.
-The 'tidyverse' (Citation) and 'opendatatoronto' (citation) packages were used to download the Victims of Crime dataset.
+The 'tidyverse' (Citation) and 'opendatatoronto' [@opendatatoronto] packages were used to download the Victims of Crime dataset.
 Then, the 'tidyverse' package (citation) was to clean and test the raw dataset.
@@
 
-After loading the dataset using the R programming language (citation), I used the 'tidyverse' (citation) package, and the 'ggplot2' (citation) package.
+After loading the dataset using the R programming language [@citeR], I used the 'tidyverse' (citation) package, and the 'ggplot2' (citation) package.
 The R code to do so was adapted from Alexander (insert citation).
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-## Data Attribution Statement {#sec-appendix-attributiojn}
+## Data Attribution Statement {#sec-appendix-attribution}
 
@@
 
-# References (To be added) {#sec-references}
+# References {#sec-references}

### 2024-09-26 paper/paper.qmd
updated title and added Table 1 in appendix

@@
 ---
-title: "Analysis of Crime Victims by Type and Gender from 2014 to 2023"
+title: "Who are Toronto’s Crime Victims? A Gender and Crime Type Analysis from 2014 to 2023"
 author: 
@@
 library(scales)
+library(knitr)
 
@@
 
+## Data Sample {#sec-appendix-sample}
+
+```{r}
+#| label: tbl-head
+#| tbl-cap: First 5 rows of cleaned data 
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Table Example ####
+cleaned_data |>
+  slice(1:5) |>
+  kable(
+    digits = 1,
+    booktabs = TRUE,
+    linesep = ""
+  )
+
+```
+
 \newpage

### 2024-09-26 paper/paper.qmd
Adjusted figures and tables for readability

@@
   geom_text(aes(label = Victims_of_Crimes),
-            position = position_dodge(width = 0.9),
+            position = position_dodge(width = 0.95),
             vjust = -0.2,
@@
   geom_text(aes(label = Victims_of_Crimes),
-            position = position_dodge(width = 0.9),
+            position = position_dodge(width = 1.05),
             vjust = -0.2,
@@
 #| label: tbl-head
-#| tbl-cap: First 5 rows of cleaned data 
+#| tbl-cap: First 5 Rows of Cleaned Data 
 #| echo: false

### 2024-09-27 scripts/00-simulate_data.R
Used lintr to clean code

@@ -29,18 +29,18 @@ data <-
     category = "Crimes Against the Person",
     
     # Randomly pick subtype
-    subtype = sample(c("Assault", "Theft", "Burglary", "Other"), size = 1244, 
-                       replace = TRUE),
+    subtype = sample(c("Assault", "Theft", "Burglary", "Other"), size = 1244,
+                     replace = TRUE),
     
     # Randomly assign assault_subtype and N/A to other crimes
     assault_subtype = ifelse(
       subtype == "Assault", 
-      sample( c("Assault Peace Officer", 
+      sample(c("Assault Peace Officer", 
                 "Assault Resist Arrest", 
                 "Assault Peace Officer Weapon/Bodily Harm", 
                 "Aggravated Peace Officer"),
-              size = sum(subtype == "Assault"), 
-              replace = TRUE),"N/A"),
+              size = sum(subtype == "Assault"),  
+              replace = TRUE),"N/A"), 
       
     # Randomly assign sex 
     sex = sample(c("M", "F", "U"), size = 1244, replace = TRUE), 

### 2024-09-27 scripts/01-download_data.R

@@ -14,20 +14,19 @@ library(tidyverse)
 
 
 #### Download data ####
+# Code is adapted from OpenDataToronto
 # get package
 package <- show_package("police-annual-statistical-report-victims-of-crime")
 
 # get all resources for this package
 resources <- list_package_resources("police-annual-statistical-report-victims-of-crime")
 
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# identify Datasource Resources
+datastore_resources <- filter(resources, tolower(format) %in% c("csv", "geojson"))
 
 # load the first datastore resource as a sample
 data <- filter(datastore_resources, row_number()==1) %>% 
   get_resource()
-
-
 
 #### Save data ####
 write_csv(data, "data/raw_data/raw_data.csv") 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 22 September 2024
 # Contact: kevin.roe@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Make sure to download the dataset
+# Pre-requisites: Run 01-download_data.R
 # Any other information needed? None.
 
 #### Workspace setup ####
@@ -18,16 +18,16 @@ cleaned_data <-
   raw_data |>
   
   # Only include the columns of interest 
-  select('_id','REPORT_YEAR','SUBTYPE','ASSAULT_SUBTYPE','SEX',
-         'COUNT_') |>
+  select("_id", "REPORT_YEAR", "SUBTYPE", "ASSAULT_SUBTYPE", "SEX",
+         "COUNT_") |>
   
   # Rename column headers for clarity
-  rename('id' = '_id',
-         'Year' = 'REPORT_YEAR',
-         'CrimeType' = 'SUBTYPE',
-         'AssaultType' = 'ASSAULT_SUBTYPE',
-         'Sex' = 'SEX',
-         'Count' = 'COUNT_') |>
+  rename("id" = "_id",
+         "Year" = "REPORT_YEAR",
+         "CrimeType" = "SUBTYPE",
+         "AssaultType" = "ASSAULT_SUBTYPE",
+         "Sex" = "SEX",
+         "Count" = "COUNT_") |>
   
 # Rename Sex for clarity purposes
   mutate(Sex = case_when(
@@ -53,7 +53,5 @@ mutate(AssaultType = case_when(
 #### Suppress the column type messages ####
 options(tibble.show_col_types = FALSE)
 
-
-  
 #### Save the Cleaned Dataset ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-27 scripts/03-test_data.R

@@ -16,28 +16,28 @@ data <- read_csv("data/analysis_data/analysis_data.csv")
 
 # Check the class of the columns
 
-data$'CrimeType' |> class() == "character"
+data$"CrimeType" |> class() == "character"
 
-data$'AssaultType' |> class() == "character"
+data$"AssaultType" |> class() == "character"
 
-data$'Sex' |> class() == "character"
+data$"Sex" |> class() == "character"
 
-data$'AgeGroup' |> class() == "character"
+data$"AgeGroup" |> class() == "character"
 
-data$'Count' |> class() == "numeric"
+data$"Count" |> class() == "numeric"
 
 
 # Test for negative numbers
-data$'Count' |> min() <= 0
+data$"Count" |> min() <= 0
 
 # Test for NAs
-all(is.na(data$'Count'))
-all(is.na(data$'AssaultType'))
+all(is.na(data$"Count"))
+all(is.na(data$"AssaultType"))
 
 # Check that the only Sex types listed are  Male, Female, and Unknown
-data$'Sex' |> 
+data$"Sex" |> 
   unique() |>
-sort() ==  sort(c("Male", "Female", "Unknown"))
+  sort() ==  sort(c("Male", "Female", "Unknown"))
 
 
 

### 2024-09-27 paper/paper.qmd
updated citations and refined abstract

@@
 date-format: long
-abstract: "This paper analyzes data from the Police Annual Statistical Report – Victims of Crime for the years 2014 to 2023 in the City of Toronto.
-The paper reveals there are 2% more male victims overall, but women are disproportionately affected in certain crime categories: female victims outnumber males by 78 percentage points in sexual crimes and by 9 percentage points in other crimes.
-These differences show how vulnerability to specific crimes varies by sex.
-Further research should focus on victims of fatal crimes to explore these disparities."
+abstract: "This paper analyzes data from the Toronto Police Service's Annual Statistical Reporton Crime Victims from 2014 to 2023, focusing on the relationship between sex and crime victimization in the city.
+The paper reveals there are 2% more male victims overall, but women are disproportionately affected in certain crime categories: female victims outnumber males by 78% in sexual crimes and by 9% in other crimes.
+These findings highlight how vulnerability to certain crimes is influenced by sex, underscoring the need for targeted policy interventions to strengthen victim support services.
+Further research should focus on victims of fatal crimes, particularly those linked to Intimate Partner Violence (IPV)."
 format: pdf
@@
 
-# Purpose: Generates a paper titled "Analysis of Crime Victims by Type and gender from 2014 to 2023" in PDF format when rendered.
+# Purpose: Generates a paper titled "Who are Toronto’s Crime Victims? A Gender and Crime Type Analysis from 2014 to 2023" in PDF format when rendered.
 # Author: Kevin Roe
@@
 #install.packages("tidyverse")
-# You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.
-# The remainder of this paper is structured as follows. @sec-mydatasection
+#install.packages("ggplot2")
+#install.packages("here")
+#install.packages("knitr)
+
 library(tidyverse)
+library(here)
 library(ggplot2)
@@
 Notably, certain demographic groups have been more affected by crime than others.
-For example, in the United States in 2008, there were a greater number of female nonfatal violent domestic violence victims than male ones (Catalano et al.
-INSERT STATS).
+For example, in the United States in 2008, there were a greater number of female nonfatal violent domestic violence victims than male ones [@dojfemalevictims].
 Similarly, Toronto is no exception to this phenomenon.
 The City of Toronto notes that certain groups of people are more susceptible to Gender-based violence than others.
+
+Similarly, Toronto is no exception to this.
+The City of Toronto notes that certain groups of people are more susceptible to Gender-based violence than others [@toronto_gbv].
+While the relationship between one’s sex and crime has been studied at the national level, there have been no studies or reports specific to Toronto, leading the paper to explore this gap.
+For Toronto to develop effective policies and build robust victim support and recovery, it is important to analyze the disparities in the types of crimes the different sexes face [@canada_family_violence].
 
@@
 
-The remainder of the paper is structured as follows: @sec-data-overview discusses the data; @sec-data-results presents the results; and @sec-discussion discusses our findings and highlights key weaknesses in the study.
+The remainder of the paper is structured as follows: @sec-data-overview discusses the data and the relevant measurement methodologies; @sec-data-results presents the data's results; and @sec-discussion discusses our findings, highlights the implications of them, and notes any key weaknesses in the study.
+
+\newpage
 
@@
 
-This dataset, “Police Annual Statistical Report – Victims of Crime”, was published and refreshed on August 2nd, 2024, by the Toronto Police Services.
-The Toronto Police Service publishes various datasets on public safety to inform the public with information regarding public safety and self-awareness.
+This dataset, “Police Annual Statistical Report – Victims of Crime”, was published and refreshed on August 2nd, 2024, by the Toronto Police Services [@victimsofcrime].
+The Toronto Police Service publishes various datasets on public safety to inform the public with information regarding public safety and self-awareness [@annual_statistics_report].
+Data on the victims of crime is part of the overall Toronto Police Service’s Annual Statistic Report and also reports on reported crimes, victims of crimes, search of persons, firearms, and the Police Service’s budget [@victimsofcrime].
+The data is collected using historical arrest data in the City of Toronto to report on the victims age, reported year, gender, and their crime then counts the number of people who fit these categories into one entry.The Toronto Police Service publishes various datasets on public safety to inform the public with information regarding public safety and self-awareness [@victimsofcrime].
 Following the Municipal Freedom of Information and Protection of Privacy Act, the Toronto Police Service ensured to also protect the privacy of individuals involved in the reported crimes.
-The dataset is updated annually, is open data, and can be used if an attribution statement @sec-appendix-attribution and is properly cited (Insert Citation).
+The dataset is updated annually, is open data, and can be used if an attribution statement @sec-appendix-attribution and is properly cited [@tphlicense].
 
-The variables of interest in the paper are CrimeType (named “CATEGORY” in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robberty and Other; AssaultType (named “ASSAULT_SUBTYPE” in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named “SEX” in the original dataset), which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count, which notes the number of identified victims who share the same demographic characteristics reported in the dataset.
+The variables of interest in the paper are CrimeType (named SUBTYPE in the original dataset, which categories the type of crime into the four areas of Assault, Sexual Violation, Robbery and Other; AssaultType (named ASSAULT_SUBTYPE in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named SEX in the original dataset), which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count (named COUNT in the original dataset), which counts the number of identified victims who share the same demographic characteristics previously.
+Each entry in the dataset does not represent a unique person but notes the numbre of people who share the same characteristics, such as Sex and CrimeType.
 Other characteristics such as age group, reported year and age cohort were not included as the variables were beyond the scope of the paper.
+Using knitr [@citeknitr], the first 10 lines of the dataset used for analysis is shown in @sec-appendix-sample under @tbl-head for clarity purposes.
 
 The paper uses the R programming language [@citeR] to analyze the dataset.
-The 'tidyverse' package was used to simulate the dataset and generate tests.
-The 'tidyverse' (Citation) and 'opendatatoronto' [@opendatatoronto] packages were used to download the Victims of Crime dataset.
-Then, the 'tidyverse' package (citation) was to clean and test the raw dataset.
-Finally, 'ggplot2' (citation), 'tidyverse' (citation), and 'scales' (citation) packages were used to create data visualization figures to highlight the results.
+The tidyverse package was used to simulate the dataset and generate tests.
+The tidyverse [@citetidyverse] and opendatatoronto [@citeopendatatoronto] packages were used to download the Victims of Crime dataset.
+Then, the tidyverse[@citetidyverse] package was to clean and test the raw dataset.
+Finally, ggplot2 [@citeggplot2], tidyverse [@citetidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to highlight the results.
+\newpage
 
@@
 
-After loading the dataset using the R programming language [@citeR], I used the 'tidyverse' (citation) package, and the 'ggplot2' (citation) package.
-The R code to do so was adapted from Alexander (insert citation).
+After loading the dataset using the R programming language [@citeR] and the here package [@citehere], I used the tidyverse [@citetidyverse] package, and the ggplot2 [@citeggplot2] package.
+The R code to do so was adapted from @tellingstorieswithdata.
 
@@
 @fig-crimetype shows that while there are more assault and robbery male victims, the number of female victims for sexual crimes and other crimes remain significantly higher than male ones.
-Female victims outnumber males by 78 percentage points in sexual crimes and by 9 percentage points in other crimes.
+Female victims outnumber males by 78\% in sexual crimes and by around 9\% in other crimes.
 
@@
 @fig-assault shows that diving deeper into the different crimes under assault shows that there are more male assault victims than female ones  across all types of assault, including different types of assault to a peace officer.
+The specified assault categories are crimes to police officers and the Other subcategory is the more prevalent one compared to the others.
 
@@
 
+## Main Findings {#sec-discussion-findings}
+
 @fig-total showed that there were 2.1\% more reported male crime victims than female ones.
-However, when we break this down, the number of female victims outnumber males by 78 percentage points and 9 percentage points for sexual and other crimes, respectively.
+However, when we break this down, the number of female victims outnumber males by 78\% and 9\% for sexual and other crimes, respectively.
 Further, when dissecting assault victims, we don’t see this same differential and there are more male victims than female ones.
@@
 @fig-crimetype shows that women are more susceptible to gender-based violence such as sexual assault or harassment.
-According to Statistics Canada, approximately 4.7 million or 30\% of all women aged 15 or older have been sexually assaulted outside of an intimate relationship at least once.
-The data in figure 2 shows that from 2014 to 2023 there were 21,515 instances of sexual violence to women in the city of Toronto and only 3,181 for men, a disproportionate difference.
+According to Cotter, approximately 4.7 million or 30\% of all women aged 15 or older have been sexually assaulted outside of an intimate relationship at least once [@ipv_statcan].
+The data in figure 2 shows that, from 2014 to 2023, there were 21,515 instances of sexual violence to women in the city of Toronto and only 3,181 for men, a disproportionate difference.
 
@@
 @fig-crimetype shows that, across the past nine years, there were 11,169 more male assault victims than female.
-When broken into different categories, @fig-assault highlights that there are more male victims across all subsections of assault than female.
+When broken into different categories, @fig-assault highlights that there are more male victims across all subsections of assault than female.@fig-assault also shows that most crimes in assault are not to police officers but fall under a more broad group of assault charges.
+Figure 3 is particularly highlighted because the dataset is generated by the Toronto Police Department, and for the Department, it is useful to highlight how many cases of assault were on on-duty police officers.
 Because assault is the most common category of crime in Toronto, the 2\% difference between male and female victims in @fig-total is fueled by the difference underlined in the assault category.
@@
 
+## Implications {#sec-discussion-implications}
+
+The result of this study has numerous implications in policy making, victim support, and overall social justice initiatives.
+Most importantly, by highlighting these differences, the study highlights specific vulnerabilities, such as women to sexual crimes or men to assault.
+In particular, the results are important to municipal policy makers designing legal protections, interventions and support services, which is important for aiding victims [@coe_gender_violence].
+Further, understanding the specific crimes that affect different sexes will empower victim support services.
+While further study is needed to evaluate the different sexes’ responses to crimes, the study still highlights these differences in the meaningful way that is specific to the Toronto community.
+Finally, the result of the study opens up broader discussions on systemic issues like misogyny and toxic gender roles that make discussion on the higher crime rates against specific sexes possible [@canadian_womens_foundation].
+Overall, there are various legislative and social implications of the study.
+
+## Limitations and Next Steps {#sec-discussion-limitations}
+
 When working with the dataset, there are a few limitations and next steps to keep in mind.
-First, the data may have some duplicates because the "Count"(INSERT CHANGE) variable could include the same person if they were victimized for the same offence during the same period.
-In addition, the dataset does not breakdown other into specific crimes under "CrimeType", making it difficult to analyze fatal victims of crime.
-Further research should look at if women or men are more susceptible to fatal crimes such as manslaughter or first-degree murder.
+First, the data may have some duplicates because the Count could include the same person if they were victimized for the same offence during the same period.
+In addition, the dataset does not breakdown other into specific crimes under CrimeType, making it difficult to analyze fatal victims of crime.
+Similarly, the data set does not distinguish or note the severity of crime and generalizes it into different categories and makes it difficult to show that women in Toronto are more susceptible to more fatal crimes.
+For example, research by the Government of Canada suggests that women are more susceptible to Intimate Partner Violence (IPV), such as being choked, assaulted, or sexually assaulted [@statcan_gbv_2019].
+Further research should investigate not only fatal crimes such as manslaughter and murder, but should also investigate if women in Toronto are more susceptible to IPV than men, a key area in gender-based violence.
 
@@
 
-"This data contains information licensed under the Open Government License - Toronto" (insert citation).
+"This data contains information licensed under the Open Government License - Toronto" [@tphlicense].
 

### 2024-09-27 paper/paper.qmd
Proofread paper for final submission and added summary statistics

@@
 date-format: long
-abstract: "This paper analyzes data from the Toronto Police Service's Annual Statistical Reporton Crime Victims from 2014 to 2023, focusing on the relationship between sex and crime victimization in the city.
-The paper reveals there are 2% more male victims overall, but women are disproportionately affected in certain crime categories: female victims outnumber males by 78% in sexual crimes and by 9% in other crimes.
+abstract: "This paper analyzes data from the Toronto Police Service's Annual Statistical Report on Crime Victims from 2014 to 2023, focusing on the relationship between sex and crime victimization in the city.
+Overall, male victims outnumber female victims by 2%, but women are disproportionately affected in certain crime categories: female victims outnumber males by 78% in sexual crimes and by 9% in other crimes.
 These findings highlight how vulnerability to certain crimes is influenced by sex, underscoring the need for targeted policy interventions to strengthen victim support services.
@@
 Notably, certain demographic groups have been more affected by crime than others.
-For example, in the United States in 2008, there were a greater number of female nonfatal violent domestic violence victims than male ones [@dojfemalevictims].
-Similarly, Toronto is no exception to this phenomenon.
-The City of Toronto notes that certain groups of people are more susceptible to Gender-based violence than others.
+For example, in the United States in 2008, there were a greater number of female non-fatal  domestic violence victims than male ones [@dojfemalevictims].
 
@@
 The City of Toronto notes that certain groups of people are more susceptible to Gender-based violence than others [@toronto_gbv].
-While the relationship between one’s sex and crime has been studied at the national level, there have been no studies or reports specific to Toronto, leading the paper to explore this gap.
+While the relationship between sex and crime victimization has been studied at the federal level, there have been no studies or reports specific to Toronto, leading the paper to explore this gap.
 For Toronto to develop effective policies and build robust victim support and recovery, it is important to analyze the disparities in the types of crimes the different sexes face [@canada_family_violence].
 
-In this paper, I analyze the number of crime victims by sex in Toronto from 2014 to 2023.
+In this paper, I will analyze the number of crime victims by sex in Toronto from 2014 to 2023.
 I categorize crime into four different areas: Assault, Sexual Violation, Robbery, and Other.
-Furthermore, assault is further broken down into different types of assault noted by the Toronto Police Department: assault on a peace officer, aggravated assault on a peace officer, resisting arrest, and assault on a peace officer with a weapon.
-We find that there are more female victims of sexual crimes and other broad crimes than men, there are more male victims for non-fatal crimes such as assault or robbery.
+Furthermore, the Assault category is further broken down into different types of assault noted by the Toronto Police Department: Assault on a Peace Officer, Aggravated Assault on a Peace Officer, Resisting Arrest, and Assault on a Peace Officer with a Weapon.
+Our findings indicate that while there are more female victims of sexual crimes and other broad categories of crime, the overall number of male victims is higher, especially in assault and robbery.
 
@@
 
-This dataset, “Police Annual Statistical Report – Victims of Crime”, was published and refreshed on August 2nd, 2024, by the Toronto Police Services [@victimsofcrime].
-The Toronto Police Service publishes various datasets on public safety to inform the public with information regarding public safety and self-awareness [@annual_statistics_report].
-Data on the victims of crime is part of the overall Toronto Police Service’s Annual Statistic Report and also reports on reported crimes, victims of crimes, search of persons, firearms, and the Police Service’s budget [@victimsofcrime].
-The data is collected using historical arrest data in the City of Toronto to report on the victims age, reported year, gender, and their crime then counts the number of people who fit these categories into one entry.The Toronto Police Service publishes various datasets on public safety to inform the public with information regarding public safety and self-awareness [@victimsofcrime].
-Following the Municipal Freedom of Information and Protection of Privacy Act, the Toronto Police Service ensured to also protect the privacy of individuals involved in the reported crimes.
+This dataset, “Police Annual Statistical Report – Victims of Crime”, was published and refreshed on August 2nd, 2024, by the Toronto Police Service [@victimsofcrime].
+The Toronto Police Service publishes various datasets on public safety to inform the public about safety issues [@annual_statistics_report].
+Data on crime victims is included in the Toronto Police Service’s Annual Statistical Report, which also covers reported crimes, searches of persons, firearms, and the Police Service’s budget [@annual_statistics_report].
+The data is collected using historical arrest data in the City of Toronto to report on the victim's age, reported year, gender, and the type of crime.
+It then counts the number of people who fit these categories into a single entry.
+Following the Municipal Freedom of Information and Protection of Privacy Act, the Toronto Police Service ensures to protect the privacy of individuals involved in the reported crimes when publishing the data.
 The dataset is updated annually, is open data, and can be used if an attribution statement @sec-appendix-attribution and is properly cited [@tphlicense].
 
-The variables of interest in the paper are CrimeType (named SUBTYPE in the original dataset, which categories the type of crime into the four areas of Assault, Sexual Violation, Robbery and Other; AssaultType (named ASSAULT_SUBTYPE in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named SEX in the original dataset), which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count (named COUNT in the original dataset), which counts the number of identified victims who share the same demographic characteristics previously.
-Each entry in the dataset does not represent a unique person but notes the numbre of people who share the same characteristics, such as Sex and CrimeType.
-Other characteristics such as age group, reported year and age cohort were not included as the variables were beyond the scope of the paper.
-Using knitr [@citeknitr], the first 10 lines of the dataset used for analysis is shown in @sec-appendix-sample under @tbl-head for clarity purposes.
+The variables of interest in the paper are CrimeType (named SUBTYPE in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robbery and Other; AssaultType (named ASSAULT_SUBTYPE in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named SEX in the original dataset) - which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count (named COUNT in the original dataset), which counts the number of identified victims who share the same demographic characteristics previously.
+Each entry in the dataset does not represent a unique person but notes the number of people who share the same characteristics, such as Sex and CrimeType.
+Other characteristics such as age group, reported year and age cohort were not variables of interest for the study and were removed from the cleaned dataset.
+Using knitr [@citeknitr], the first 10 lines of the cleaned dataset is shown in @sec-appendix-sample under @tbl-head.
+In addition, summaries of the Sex and Year variables are displayed in @sec-appendix-summary_statistics.
 
 The paper uses the R programming language [@citeR] to analyze the dataset.
-The tidyverse package was used to simulate the dataset and generate tests.
-The tidyverse [@citetidyverse] and opendatatoronto [@citeopendatatoronto] packages were used to download the Victims of Crime dataset.
-Then, the tidyverse[@citetidyverse] package was to clean and test the raw dataset.
-Finally, ggplot2 [@citeggplot2], tidyverse [@citetidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to highlight the results.
+The tidyverse package was used to simulate the dataset.
+Also, the tidyverse [@citetidyverse] and opendatatoronto [@citeopendatatoronto] packages were used to download the Victims of Crime dataset.
+Then, the tidyverse [@citetidyverse] package was used to clean the raw dataset and generate tests.
+Finally, ggplot2 [@citeggplot2], tidyverse [@citetidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to display the results.
+
 \newpage
@@
 
-After loading the dataset using the R programming language [@citeR] and the here package [@citehere], I used the tidyverse [@citetidyverse] package, and the ggplot2 [@citeggplot2] package.
+After loading the dataset using the R programming language [@citeR] and the here package [@citehere], I used the tidyverse [@citetidyverse] package, and the ggplot2 [@citeggplot2] package to summarize our results.
 The R code to do so was adapted from @tellingstorieswithdata.
@@
 
-@fig-total shows that there are 137,527 male victims of crime from 2014 to 2023 and 134726 female ones, representing  a difference of 2,801 people or 2.1\%.
-@fig-total shows that there are more male crime victims than female ones, albeit the difference does not seem to be economically significant.
+@fig-total shows that there are 137,527 male crime victims from 2014 to 2023 and 134,726 female ones, representing a difference of 2,801 people or 2.1\%.
 
@@
 ```
-@fig-assault shows that diving deeper into the different crimes under assault shows that there are more male assault victims than female ones  across all types of assault, including different types of assault to a peace officer.
-The specified assault categories are crimes to police officers and the Other subcategory is the more prevalent one compared to the others.
+@fig-assault shows that even after breaking assault category into multiple subcategories, there are more male assault victims than female ones across all subcategories of assault.
+The specified assault subcategories are specific crimes against police officers and the Other subcategory is the most prevalent subcategory.
 
@@
 
-@fig-total showed that there were 2.1\% more reported male crime victims than female ones.
+@fig-total showed that there were 2.1\% more total reported male crime victims than female ones.
 However, when we break this down, the number of female victims outnumber males by 78\% and 9\% for sexual and other crimes, respectively.
-Further, when dissecting assault victims, we don’t see this same differential and there are more male victims than female ones.
+Further, when dissecting assault victims, male victims outnumber female ones across all specified assault subcategories.
 
-Analyzing the data shows that vulnerability to specific crimes varies by sex.
-@fig-crimetype shows that women are more susceptible to gender-based violence such as sexual assault or harassment.
-According to Cotter, approximately 4.7 million or 30\% of all women aged 15 or older have been sexually assaulted outside of an intimate relationship at least once [@ipv_statcan].
-The data in figure 2 shows that, from 2014 to 2023, there were 21,515 instances of sexual violence to women in the city of Toronto and only 3,181 for men, a disproportionate difference.
+Analyzing the data underscores that vulnerability to specific crimes varies by sex.
+@fig-crimetype underscores that women are more susceptible to gender-based violence such as sexual assault or harassment.
+The paper's findings are corroborated by Cotter at the national level, where he highlights that approximately 4.7 million, or 30\%, of all women aged 15 or older have been sexually assaulted outside of an intimate relationship at least once [@ipv_statcan].
+The data in @fig-crimetype shows that, from 2014 to 2023, there were 21,515 instances of sexual violence to women in the city of Toronto and only 3,181 for men, a disproportionate difference.
 
-However, @fig-crimetype shows that men are also more susceptible to being victims of assault, and robbery  compared to women.
-@fig-crimetype shows that, across the past nine years, there were 11,169 more male assault victims than female.
-When broken into different categories, @fig-assault highlights that there are more male victims across all subsections of assault than female.@fig-assault also shows that most crimes in assault are not to police officers but fall under a more broad group of assault charges.
-Figure 3 is particularly highlighted because the dataset is generated by the Toronto Police Department, and for the Department, it is useful to highlight how many cases of assault were on on-duty police officers.
+However, @fig-crimetype suggests that men are also more susceptible to being victims of assault and robbery compared to women, as there were 11,169 more male assault victims than female one.
+In addition.
+when broken into different categories, @fig-assault highlights that there are more male victims across all subsections of assault than female victim.
+@fig-assault also notes that most crimes in assault are not to police officers but fall under a more broad group of assault charges.
 Because assault is the most common category of crime in Toronto, the 2\% difference between male and female victims in @fig-total is fueled by the difference underlined in the assault category.
@@
 The result of this study has numerous implications in policy making, victim support, and overall social justice initiatives.
-Most importantly, by highlighting these differences, the study highlights specific vulnerabilities, such as women to sexual crimes or men to assault.
+Most importantly, the results highlights the differences in crime victimization between sexes, such as women to sexual crimes or men to assault.
 In particular, the results are important to municipal policy makers designing legal protections, interventions and support services, which is important for aiding victims [@coe_gender_violence].
-Further, understanding the specific crimes that affect different sexes will empower victim support services.
+Further, understanding the specific crimes that affect different sexes will empower victim support services in designing targeted and effective.
 While further study is needed to evaluate the different sexes’ responses to crimes, the study still highlights these differences in the meaningful way that is specific to the Toronto community.
-Finally, the result of the study opens up broader discussions on systemic issues like misogyny and toxic gender roles that make discussion on the higher crime rates against specific sexes possible [@canadian_womens_foundation].
-Overall, there are various legislative and social implications of the study.
+Ultimately, the study's findings prompt broader discussion about systemic issues such as misogyny and toxic gender roles, as well as their effects on the community [@canadian_womens_foundation].
+Overall, there are various legislative and social implications of the study that are relevant to Toronto.
 
@@
 First, the data may have some duplicates because the Count could include the same person if they were victimized for the same offence during the same period.
-In addition, the dataset does not breakdown other into specific crimes under CrimeType, making it difficult to analyze fatal victims of crime.
-Similarly, the data set does not distinguish or note the severity of crime and generalizes it into different categories and makes it difficult to show that women in Toronto are more susceptible to more fatal crimes.
+In addition, the dataset does not breakdown the Other cateogry into specific crimes, making it difficult to analyze fatal victims of crime.
+Similarly, the data set does not distinguish or note the severity of crime, generalizing them into broad categories, which makes it difficult to show that women in Toronto are more susceptible to more fatal crimes.
 For example, research by the Government of Canada suggests that women are more susceptible to Intimate Partner Violence (IPV), such as being choked, assaulted, or sexually assaulted [@statcan_gbv_2019].
-Further research should investigate not only fatal crimes such as manslaughter and murder, but should also investigate if women in Toronto are more susceptible to IPV than men, a key area in gender-based violence.
+Further research should investigate not only fatal crimes such as manslaughter and murder, but also whether women in Toronto are more vulnerable to IPV than men, a key area in gender-based violence.
 
@@
 
-#### Table Example ####
+#### Dataset Head ####
 cleaned_data |>
@@
 ```
+\newpage
+
+## Summary Statistics {#sec-appendix-summary_statistics}
+
+```{r}
+#| label: tbl-yearsummary
+#| tbl-cap: Number of Crimes per Year 
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Table of Number of Crime Victims per Year ####
+summary_data <- cleaned_data |>
+  group_by(Year) |>
+  summarise(total_count = sum(Count, na.rm = FALSE)) |>
+  ungroup()
+
+# Display data using Kable
+summary_data |> 
+  kable(
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    col.names = c("Year","Total Victims")
+  )
+
+```
+```{r}
+#| label: tbl-sexsummary
+#| tbl-cap: Number of Victims by Sex
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Table of Number of Victims by Sex ####
+summary_by_sex <- cleaned_data |>
+  group_by(Sex) |>
+  summarise(total_count = sum(Count, na.rm = TRUE)) |>
+  ungroup()
+
+# Display data using Kable
+summary_by_sex |> 
+  kable(
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    col.names = c("Sex","Total Victims")
+  )
+
+```
 

### 2024-09-27 paper/paper.qmd
Updated abstract and fixed errors in Appendix A5

@@
 Overall, male victims outnumber female victims by 2%, but women are disproportionately affected in certain crime categories: female victims outnumber males by 78% in sexual crimes and by 9% in other crimes.
-These findings highlight how vulnerability to certain crimes is influenced by sex, underscoring the need for targeted policy interventions to strengthen victim support services.
-Further research should focus on victims of fatal crimes, particularly those linked to Intimate Partner Violence (IPV)."
+These findings highlight how vulnerability to certain crimes is influenced by sex, underscoring the need for targeted policy initiatives to strengthen victim support services.
+Further research should focus on victims of fatal crimes, particularly those linked to Intimate Partner Violence (IPV) to address the limitations of the dataset."
 format: pdf
@@
 
-Select columns from the raw data set were filtered out and data entries were renamed for simplicity.
+Select columns from the raw data set were filtered out and data entries were renamed for clarity purposes.
 

### 2024-09-27 paper/paper.qmd
Updated graph headers for consistency

@@
 thanks: "All data, R code, and other files are available in the following GitHub repository: <https://github.com/Kanghyunroe/victims_of_crimes/tree/master>"
-date: today
+date: 09-27-2024
 date-format: long
@@
   labs(x = "Sex",
-       y = "Victims of Crimes") +
+       y = "Number of Victims") +
   scale_y_continuous(labels = comma) +
